The Rockerverse: Images, Packages, and Applications for Containerization with R

### 2020-01-23 manuscript.Rmd

\label{pipelines}
Third, containers are perfectly suited for **packaging and executing software pipelines** and required data.
Containers allow building complex processing pipelines that are independent from the host programming language.
Due to its original use case (see&nbsp;\nameref{introduction}), Docker has no standard mechanisms for chaining containers together; it lacks definitions and protocols for environment variables, volume mounts and/or ports that could enable transfer of input (parameters and data) and output (results) to and from containers.
Some packages, e.g., \pkg{containerit}, do provide Docker images that can be used similarly to CLIs, but their usage is cumbersome\footnote{\href{https://o2r.info/containerit/articles/container.html}{https://o2r.info/containerit/articles/container.html}}.
**\pkg{outsider}** ([https://antonellilab.github.io/outsider/](https://antonellilab.github.io/outsider/)) tackles the problem of integrating external programs into an R workflow \citep{bennett_outsider_2020}.
Installation and usage of external programs can be difficult, convoluted and even impossible if the platform is incompatible.
Therefore \pkg{outsider} uses the platform-independent Docker images to encapsulate processes in _outsider modules_.
Each outsider module has a `Dockerfile` and an R package with functions for interacting with the encapsulated tool.
Using only R functions, an end-user can install a module with the \pkg{outsider} package and then call module code to integrate a tool into their own R-based workflow seamlessly.
The \pkg{outsider} package and module manage the containers and handle the transmission of arguments and the transfer of files to and from a container.
These functionalities also allow a user to launch module code on a remote machine via SSH, expanding the potential computational scale.
Outsider modules can be hosted code-sharing services, e.g., on GitHub, and \pkg{outsider} contains discovery functions for them.
<!-- informative examples of how modules work, just as a reminder for @nuest
https://github.com/DomBennett/om..astral/blob/master/examples/astral.R
https://github.com/DomBennett/om..raxml#partitioned-dna-analysis
-->

>>>>

\label{pipelines}
Third, containers are perfectly suited for **packaging and executing software pipelines** and required data.
Containers allow building complex processing pipelines that are independent from the host programming language.
Due to its original use case (see&nbsp;\nameref{introduction}), Docker has no standard mechanisms for chaining containers together; it lacks definitions and protocols for environment variables, volume mounts and/or ports that could enable transfer of input (parameters and data) and output (results) to and from containers.
Some packages, e.g., \pkg{containerit}, do provide Docker images that can be used similarly to CLIs, but their usage is cumbersome\footnote{\href{https://o2r.info/containerit/articles/container.html}{https://o2r.info/containerit/articles/container.html}}.
**\pkg{outsider}** ([https://docs.ropensci.org/outsider/](https://docs.ropensci.org/outsider/)) tackles the problem of integrating external programs into an R workflow \citep{bennett_outsider_2020}.
Installation and usage of external programs can be difficult, convoluted and even impossible if the platform is incompatible.
Therefore \pkg{outsider} uses the platform-independent Docker images to encapsulate processes in _outsider modules_.
Each outsider module has a `Dockerfile` and an R package with functions for interacting with the encapsulated tool.
Using only R functions, an end-user can install a module with the \pkg{outsider} package and then call module code to integrate a tool into their own R-based workflow seamlessly.
The \pkg{outsider} package and module manage the containers and handle the transmission of arguments and the transfer of files to and from a container.
These functionalities also allow a user to launch module code on a remote machine via SSH, expanding the potential computational scale.
Outsider modules can be hosted code-sharing services, e.g., on GitHub, and \pkg{outsider} contains discovery functions for them.
<!-- informative examples of how modules work, just as a reminder for @nuest
https://github.com/DomBennett/om..astral/blob/master/examples/astral.R
https://github.com/DomBennett/om..raxml#partitioned-dna-analysis
-->

### 2020-01-23 manuscript.tex

\label{pipelines} Third, containers are perfectly suited for
\textbf{packaging and executing software pipelines} and required data.
Containers allow building complex processing pipelines that are
independent from the host programming language. Due to its original use
case (see~\nameref{introduction}), Docker has no standard mechanisms for
chaining containers together; it lacks definitions and protocols for
environment variables, volume mounts and/or ports that could enable
transfer of input (parameters and data) and output (results) to and from
containers. Some packages, e.g., \pkg{containerit}, do provide Docker
images that can be used similarly to CLIs, but their usage is
cumbersome\footnote{\href{https://o2r.info/containerit/articles/container.html}{https://o2r.info/containerit/articles/container.html}}.
\textbf{\pkg{outsider}} (\url{https://antonellilab.github.io/outsider/})
tackles the problem of integrating external programs into an R workflow
\citep{bennett_outsider_2020}. Installation and usage of external
programs can be difficult, convoluted and even impossible if the
platform is incompatible. Therefore \pkg{outsider} uses the
platform-independent Docker images to encapsulate processes in
\emph{outsider modules}. Each outsider module has a \texttt{Dockerfile}
and an R package with functions for interacting with the encapsulated
tool. Using only R functions, an end-user can install a module with the
\pkg{outsider} package and then call module code to integrate a tool
into their own R-based workflow seamlessly. The \pkg{outsider} package
and module manage the containers and handle the transmission of
arguments and the transfer of files to and from a container. These
functionalities also allow a user to launch module code on a remote
machine via SSH, expanding the potential computational scale. Outsider
modules can be hosted code-sharing services, e.g., on GitHub, and
\pkg{outsider} contains discovery functions for them.

@@

\address{%
Dom Bennett\\
University of Gothenburg, Sweden\\
Department of Biological and Environmental Sciences\\ Carl Skottsbergs gata 22B\\ 413 19 Gothenburg, Sweden\\ \orcid{0000-0003-2722-1359}\\
}
\href{mailto:dominic.john.bennett@gmail.com}{\nolinkurl{dominic.john.bennett@gmail.com}}

>>>>

\label{pipelines} Third, containers are perfectly suited for
\textbf{packaging and executing software pipelines} and required data.
Containers allow building complex processing pipelines that are
independent from the host programming language. Due to its original use
case (see~\nameref{introduction}), Docker has no standard mechanisms for
chaining containers together; it lacks definitions and protocols for
environment variables, volume mounts and/or ports that could enable
transfer of input (parameters and data) and output (results) to and from
containers. Some packages, e.g., \pkg{containerit}, do provide Docker
images that can be used similarly to CLIs, but their usage is
cumbersome\footnote{\href{https://o2r.info/containerit/articles/container.html}{https://o2r.info/containerit/articles/container.html}}.
\textbf{\pkg{outsider}} (\url{https://docs.ropensci.org/outsider/})
tackles the problem of integrating external programs into an R workflow
\citep{bennett_outsider_2020}. Installation and usage of external
programs can be difficult, convoluted and even impossible if the
platform is incompatible. Therefore \pkg{outsider} uses the
platform-independent Docker images to encapsulate processes in
\emph{outsider modules}. Each outsider module has a \texttt{Dockerfile}
and an R package with functions for interacting with the encapsulated
tool. Using only R functions, an end-user can install a module with the
\pkg{outsider} package and then call module code to integrate a tool
into their own R-based workflow seamlessly. The \pkg{outsider} package
and module manage the containers and handle the transmission of
arguments and the transfer of files to and from a container. These
functionalities also allow a user to launch module code on a remote
machine via SSH, expanding the potential computational scale. Outsider
modules can be hosted code-sharing services, e.g., on GitHub, and
\pkg{outsider} contains discovery functions for them.

@@

\address{%
Dom Bennett\\
Gothenburg Global Biodiversity Centre, Sweden\\
Carl Skottsbergs gata 22B\\ 413 19 Göteborg, Sweden\\ \orcid{0000-0003-2722-1359}\\
}
\href{mailto:dominic.john.bennett@gmail.com}{\nolinkurl{dominic.john.bennett@gmail.com}}
